---
title: "Maryland Poison Center Graphics"
author: "Leah Jager"
date: "6/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My goal is to make a map of Maryland by county that is colored by some count or rate from the MPC data set I have created.  I would like to make this map for each year, 2006-2018, and then animate it to show changes over time.

```{r}
library(tidyverse)
library(ggmap)
library(maps)
library(gganimate)
library(gifski)
library(transformr)  # to use polygons in animation
```

First, read in the MPC data:
```{r}
mpcData <- read_csv("MPCdataFINAL.csv")
names(mpcData)
dim(mpcData)
glimpse(mpcData)
```

Now just playing with making maps of Maryland:
```{r}
#making a plot of counties in maryland using the plotting within map package
map('county', 'maryland', fill = TRUE, col = palette())

#making a plot of counties in maryland using ggplot
md <- map_data('county', 'maryland')
dim(md)
head(md)
tail(md)

# fill is black and county lines are in white
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  theme_bw()

# no fill, red outline
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, group = group), fill="NA", color="red") + 
  coord_fixed(1.3) + 
  theme_bw()

# orange fill, black outline, go orioles?
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, group = group), fill="orange", color="black") + 
  coord_fixed(1.3) + 
  theme_bw()

# give each county a different color
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, fill=subregion, group = group), color="white") + 
  coord_fixed(1.3) + 
  theme_bw()

# give each county a different color and remove color legend
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, fill=subregion, group = group), color="white") + 
  coord_fixed(1.3) + 
  theme_bw() + 
  guides(fill=FALSE) 

# get rid of gridlines
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, fill=subregion, group = group), color="white") + 
  coord_fixed(1.3) + 
  theme_nothing() + 
  guides(fill=FALSE) 

# add state outline in black
state.md <- map_data('state', 'maryland')
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, fill=subregion, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_nothing() + 
  guides(fill=FALSE) 
```

Now going to try to color by the number of human exposures:
```{r}
mpc2018 <- mpcData %>%
  filter(Year==2018) %>%
  mutate(subregion=tolower(County))

mpc2018$subregion
unique(md$subregion)

# need to remove the , md and the apostrophes
mpc2018$subregion = str_replace_all(mpc2018$subregion,", md", "")
mpc2018$subregion = str_replace_all(mpc2018$subregion," county", "")
mpc2018$subregion = str_replace_all(mpc2018$subregion,"’", "")
mpc2018$subregion = str_replace_all(mpc2018$subregion,"[.]", "")

mpc2018$subregion
unique(md$subregion)

plotData <- inner_join(md, mpc2018, by="subregion")

dim(plotData)
dim(table(plotData$subregion))

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=TotalHumanExposures, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

```

These plots are pretty meaningless without turning them into a rate using the population, but first lets just make a fdw more:
```{r}
ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=MedOutcome_Death, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=Intentional_SuspectedSuicide, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=Age_Under12mo, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()
```

First we can consider a scale change:
```{r}
ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=Age_Under12mo, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw() + scale_fill_gradient(trans="log10")
```

Now I need to find some population data by county:
```{r}
popData <- read_csv("./PEP_2018_PEPANNRES/PEP_2018_PEPANNRES.csv")
glimpse(popData)

pop2018 <- popData %>%
  select(respop72018, `GEO.display-label`) %>%
  mutate(subregion=tolower(`GEO.display-label`))

unique(plotData$subregion)
pop2018$subregion

# need to remove the , md and the apostrophes
pop2018$subregion = str_replace_all(pop2018$subregion,", maryland", "")
pop2018$subregion = str_replace_all(pop2018$subregion," county", "")
pop2018$subregion = str_replace_all(pop2018$subregion,"'", "")
pop2018$subregion = str_replace_all(pop2018$subregion,"[.]", "")

unique(plotData$subregion)
pop2018$subregion

plotData <- inner_join(plotData, pop2018, by="subregion")
plotData <- plotData %>%
  mutate(THEperCap=TotalHumanExposures/respop72018)

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=THEperCap, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

```

Now to rescale to per 1000 persons
```{r}
plotData <- plotData %>%
  mutate(THEperCapThou=TotalHumanExposures/respop72018*1000)

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=THEperCapThou, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

```

There doesn't seem to be anything that interesting when looking at total human exposures in 2018.  Can I look at what's happening over time?  I have population data from 2010-2018, so I am going to work with this!

```{r}
mpcPlotData <- mpcData %>%
  filter(Year >= 2010) %>%
  mutate(subregion=tolower(County))

mpcPlotData$subregion
unique(md$subregion)

# need to remove the , md and the apostrophes
mpcPlotData$subregion = str_replace_all(mpcPlotData$subregion,", md", "")
mpcPlotData$subregion = str_replace_all(mpcPlotData$subregion," county", "")
mpcPlotData$subregion = str_replace_all(mpcPlotData$subregion,"’", "")
mpcPlotData$subregion = str_replace_all(mpcPlotData$subregion,"[.]", "")

mpcPlotData$subregion
unique(md$subregion)

plotData <- inner_join(md, mpcPlotData, by="subregion")

dim(plotData)
dim(table(plotData$subregion))

popData <- read_csv("./PEP_2018_PEPANNRES/PEP_2018_PEPANNRES.csv")
glimpse(popData)

popLongData <- popData %>%
  select(`2010`=respop72010, `2011`=respop72011, `2012`=respop72012, `2013`=respop72013, `2014`=respop72014, `2015`=respop72015, `2016`=respop72016, `2017`=respop72017, `2018`=respop72018, subregion=`GEO.display-label`) %>%
  mutate(subregion=tolower(subregion)) %>%
  gather(Year, Population, `2010`:`2018`) %>%
  mutate(Year=as.numeric(Year))

unique(popLongData$subregion)
unique(plotData$subregion)

# need to remove the , md and the apostrophes
popLongData$subregion = str_replace_all(popLongData$subregion,", maryland", "")
popLongData$subregion = str_replace_all(popLongData$subregion," county", "")
popLongData$subregion = str_replace_all(popLongData$subregion,"'", "")
popLongData$subregion = str_replace_all(popLongData$subregion,"[.]", "")

unique(popLongData$subregion)
unique(plotData$subregion)

unique(popLongData$Year)
unique(plotData$Year)

plotData <- inner_join(plotData, popLongData, by=c("subregion", "Year"))

plotData <- plotData %>%
  mutate(THEperCapThou=TotalHumanExposures/Population*10000)

```

First I will make the plots faceted:
```{r}
ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=THEperCapThou, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw() + facet_wrap(vars(Year))
```

Going to remove the PG and M counties, since they don't have accurate counts because some (most?) of their calls are rounted to Poison Center in DC.  Will this be straightforward to do?

```{r}
plotData <- plotData %>% 
  filter(subregion!="montgomery" & subregion!="prince georges")

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=THEperCapThou, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw() + facet_wrap(vars(Year))
```

Does there seem to be an overall trend over time?
```{r}
plotData %>%
  group_by(Year) %>%
  summarize(number=sum(TotalHumanExposures), pop=sum(Population), rate=number/pop*1000)
```

What's the range by state?
```{r}
plotData %>%
  group_by(Year) %>%
  summarize(min=min(THEperCapThou), max=max(THEperCapThou))
```

Now to animate this!  (Might not be that interesting with only 9 panels, but I want to learn how to do it!)

```{r}
ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=THEperCapThou, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  labs(title = "Total human exposures per 10,000 individuals in {closest_state}") +
  coord_fixed(1.3) + 
  theme_bw() + 
  transition_states(Year, 3, 1) +
  enter_fade() +
  exit_fade()
  
```

Now I'm just going to work on making everything pretty.  First I want to show the exposure rate changing over time for the entire state.  Then I will do it county by county.

```{r}
### MPC data
mpcData <- read_csv("MPCdataFINAL.csv")
### md county line map definition data
md <- map_data('county', 'maryland')
### md state line map definition data
state.md <- map_data('state', 'maryland')
### county population data
popData <- read_csv("./PEP_2018_PEPANNRES/PEP_2018_PEPANNRES.csv")

### filter to just 2010-2018; get subregions to match
mpcPlotData <- mpcData %>%
  filter(Year >= 2010) %>%
  mutate(subregion=tolower(County))

mpcPlotData$subregion = str_replace_all(mpcPlotData$subregion,", md", "")
mpcPlotData$subregion = str_replace_all(mpcPlotData$subregion," county", "")
mpcPlotData$subregion = str_replace_all(mpcPlotData$subregion,"’", "")
mpcPlotData$subregion = str_replace_all(mpcPlotData$subregion,"[.]", "")

# join mpc to map data
plotData <- inner_join(md, mpcPlotData, by="subregion")

# organize population data; get subregions to match
popLongData <- popData %>%
  select(`2010`=respop72010, `2011`=respop72011, `2012`=respop72012, `2013`=respop72013, `2014`=respop72014, `2015`=respop72015, `2016`=respop72016, `2017`=respop72017, `2018`=respop72018, subregion=`GEO.display-label`) %>%
  mutate(subregion=tolower(subregion)) %>%
  gather(Year, Population, `2010`:`2018`) %>%
  mutate(Year=as.numeric(Year))

popLongData$subregion = str_replace_all(popLongData$subregion,", maryland", "")
popLongData$subregion = str_replace_all(popLongData$subregion," county", "")
popLongData$subregion = str_replace_all(popLongData$subregion,"'", "")
popLongData$subregion = str_replace_all(popLongData$subregion,"[.]", "")

# join population data to mpc and map data
plotData <- inner_join(plotData, popLongData, by=c("subregion", "Year"))

# create THE per 10,000 rate variable
plotData <- plotData %>%
  mutate(theRate=TotalHumanExposures/Population*10000)

# give PG and Mont counties a NA for "theRate" variable, so they will show as grey
plotData <- plotData %>%
  mutate(theRate=ifelse(subregion=="prince georges" | subregion=="montgomery", NA,TotalHumanExposures/Population*10000))

# to get text of year to show in middle of plot
plotData <- plotData %>%
  mutate(xloc=-78.5, yloc=38.75)

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=theRate, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  scale_fill_gradient(low = "white", high = "purple", na.value="grey80") + 
  theme_void() +
  labs(title="Total human exposures per 10,000 individuals", fill="Rate per 10k", caption="Data source: (1) Maryland Poison Center, (2) U.S. Census Bureau, Population Division") + theme(legend.position=c(.92,.7), legend.text.align=0, plot.title=element_text(hjust=.15, face="bold")) + 
  geom_text(data=plotData, aes(y=yloc, x=xloc, label=as.character(Year)), check_overlap = TRUE, size=10, fontface="bold") +
  coord_fixed(1.3) + 
  transition_states(Year, 3, 20)
```


Now to make this plot code "nicer" and easier to read:

```{r}
# initial map information
baseMap <- ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=theRate, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  scale_fill_gradient(low = "white", high = "purple", na.value="grey80") +
  labs(title="Total human exposures per 10,000 individuals", fill="Rate per 10k", caption="Data sources: (1) Maryland Poison Center (2) U.S. Census Bureau, Population Division \n Prince George's and Montgomery Counties (in gray) excluded from this analysis since complete \n   data from these counties requires consideration of calls to National Capitol Poison Center.") 

# make minimal theme, change position of legend and title
upgradedMap <- baseMap +
  theme_void() +
  theme(legend.position=c(.92,.7), legend.text.align=0, plot.title=element_text(hjust=.1, face="bold"), plot.caption=element_text(hjust=0)) 

# animate and add year label to animation
animatedMap <- upgradedMap +
  geom_text(data=plotData, aes(y=yloc, x=xloc, label=as.character(Year)), check_overlap = TRUE, size=10, fontface="bold") +
  coord_fixed(1.3) + 
  transition_states(Year, 3, 20)

# save as gif
mapGIF <- animate(animatedMap) 
```
