---
title: "mpc_data_temp"
author: "kexin wang"
date: "10/9/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(pdftools)
```

## READ & EXTYRACT 

```{r}
read <- function(pdf.file){
    require(dplyr)
    require(pdftools)
  # read in the pdf document; select the first page 
  pdfData <- pdf_data(pdf.file)
  p1Data <- pdfData[[1]] 
  return(p1Data)

}

ex <- function(p1Data){
   # get the year and country from the header
  year <- p1Data %>% arrange(y,x) %>%
    slice(1) %>% select(text) %>% as.numeric()
  
  county <- p1Data %>%  group_by(y) %>%
  arrange(x, .by_group=TRUE) %>% 
  summarize(line = paste(text, collapse=" ")) %>%
  slice(2) %>% select(line) %>% as.character()
  
  return(list(county, year))
}
#temp <- read(d_AG)
#ex(temp)[[1]]
```
- arrange() sort a variable in descending order
- n() count value 


## SLICE
```{r}
mod <- function(p1Data){
  
  county <- ex(p1Data)[[1]]
  year <- ex(p1Data)[[2]]
  
  cut <- p1Data %>%group_by(y)%>% 
    arrange(y) %>% summarize(n=n()) 
  
  if (county=="Prince Georges County, MD" | county=="Montgomery County, MD") {
  y.cut <- cut %>%
    slice(6) %>% select(y) %>% as.numeric()   # select row by position
  } 
  else {
    y.cut <- cut %>%
      slice(3) %>% select(y) %>% as.numeric()
  }
  
  p1Data <- p1Data %>% 
    filter(y > y.cut + 1) %>%
    mutate(column=ifelse(x < 265, "Left", "Right"))
  # create the column variable (Left/Right) since original pdf has two parts
  return(p1Data)
}
```


### try on file

```{r}
t <- read("Prince Georges County Statistical Report 2018.pdf")
#t <- read((paste0("./data/",files[24])))
ex(t)

t1 <- mod(t)
t1
```

## GROUP

```{r}
gp <- function(p1Data){
  # group the data by column and height on the page
# keep the last entry of that column/height as the value
# assign the remaining entries for that column/height the name
  data <- p1Data %>% 
    group_by(column,y) %>% 
    arrange(x, .by_group=TRUE) %>%
     # will sort first by grouping variable. Applies to grouped data frames only.
    mutate(type = ifelse(x==max(x), 
                         "value", "name")) %>%
    summarize(variable = paste(text[type=="name"], 
                               collapse=" "), 
              count=text[type=="value"]) %>% 
    filter(count != "Calls", count!="exposure", count!="Site", count!="Outcome")
  
  return(data)
  
}

comb <- function(data, year, county){
  # create the data frame for this county/date
  myRow <- as.data.frame(t(as.numeric(data$count)))
  names(myRow) <- data$variable
  
  myRow$Year <- year
  myRow$County <- county
  return(myRow)
}
```

### try on file

```{r}
t1 %>% group_by(column,y) %>%
 arrange(x, .by_group = TRUE) %>%
  mutate(type = ifelse(x==max(x) & x==min(x), "name", 
                         ifelse(x==max(x), "value", "name"))) 
```
'.by_group = TRUE' will sort first by grouping variable. Applies to grouped data frames only.

```{r}
t2 <-  t1 %>% group_by(column,y) %>%
 arrange(x, .by_group = TRUE) %>%
    #mutate(type = ifelse(x==max(x), 
     #                    "value", "name")) %>%
  mutate(type = ifelse(x==max(x) & x==min(x), "name", 
                         ifelse(x==max(x), "value", "name"))) %>%
    summarize(variable = paste(text[type=="name"], 
                               collapse=" "), 
              count=text[type=="value"])  %>%   # summarize by row?
    #filter(count != c("Calls","exposure","Site","Outcome"))
  filter(count != "Calls", count!="exposure", count!="Site", count!="Outcome")


t2
```

```{r, warning=FALSE}

fna_1 <- function(file){
  p1Data <- read(file)
  county <- ex(p1Data)[[1]]
  year <- ex(p1Data)[[2]]
  
  d <- mod(p1Data)
  d <- gp(d)
  d <- comb(d, year, county)
  return(d)
}

```



```{r}
fna_1("Prince Georges County Statistical Report 2018.pdf")
fna_1("Allegany County Statistical Report 2018.pdf")
fna_1("Talbot County Statistical Report 2018.pdf")
```


```{r}
files = list.files(path='./data',pattern="*.pdf")
```

```{r}
d1 <- fna_1(paste0("./data/",files[1]))
Data <- d1

options(warn=2)
for (i in 2:length(files)) {
  di <- fna_1(paste0("./data/",files[i]))
  Data <- bind_rows(Data,di)
  print(i)
}
```

```{r}
files[c(40,64,93)]
files[287:289]
```
```{r}
t(fna_1(paste0("./data/",files[4])))
t(fna_1(paste0("./data/",files[64])))
```

Notes:

(1) Had to modify to remove commas from counts: 5,321 to 5321 (Baltimore County 2006-[40])
(2) files[60] has wired "medical outcome" arrangement "Calvert County,2017"
(3) Have "Maryland Poison Center" printed at bottom of page 1, need to remove this (Cecil 2007-[93])
(4) I think because of the apostrophe maybe, it's saying part of the header is actually two different heights, so it's split it into 2 pieces, which means I can't cut after 3rd line, really need to cut after 4th. Make a special case for this one file.(Queen Annes 2011)
(5) Same issue as for Queen Annes, just included it in the exception  (St Marys 2011)
(6) Had to modify to account for blank counts; for now I am setting those to 0.  Should probably set them to NA? This also required me to modify how labeling the name and the value, since this row only had a name not a value. (Somerset 2006)


Here I want to try directly find keywords "Calls"

```{r}
t <- read(paste0("./data/",files[249]))
t %>%group_by(y)%>% arrange(y) 
cut <- min(t$y[t$text=="Calls"])
t1 <- t %>% filter(y > cut )
```

find the where "calls" appear at the first time, that's why I use min().

```{r update cut function}
mod <- function(p1Data){
  
  y.cut <- min(p1Data$y[p1Data$text=="Calls"])
  
  p1Data <- p1Data %>% filter(y > y.cut + 1)%>%
   mutate(column=ifelse(x < 265, "Left", "Right"))
  # create the column variable (Left/Right)
  return(p1Data)
}
```


```{r update gp}
gp <- function(p1Data){
  # group the data by column and height on the page
# keep the last entry of that column/height as the value
# assign the remaining entries for that column/height the name
  gd <- p1Data %>% 
    group_by(column,y) %>% 
    arrange(x, .by_group=TRUE) %>%
    mutate(type = ifelse(x==max(x) & x==min(x), "name", 
                         ifelse(x==max(x), "value", "name"))) 
  
    cd <- gd %>% 
      summarize(variable = paste(text[type=="name"], collapse=" "), 
              count = ifelse(is_empty(text[type=="value"])==FALSE, 
                           text[type=="value"],"0")) %>%
    filter(count != "Calls", count!="exposure", count!="Site",
           count!="Outcome",count!="Center",variable!="Maryland")
  
  return(cd)
  
}

comb <- function(data, year, county){
  # create the data frame for this county/date
  myRow <- as.data.frame(t(as.numeric(gsub(",","",data$count))))
  names(myRow) <- data$variable
  
  myRow$Year <- year
  myRow$County <- county
  return(myRow)
}
```

*Sheena* 
1.how to find empty value? is_empty()?
2.Does function max() find the maximum of x within the same group_by(y)?
3.If we can assign "name" and "value" type to these two columns respectively? 

```{r}
fna_2 <- function(file){
  p1Data <- read(file)
  county <- ex(p1Data)[[1]]
  year <- ex(p1Data)[[2]]
  d <- mod(p1Data)
  d <- gp(d)
  d <- comb(d, year, county)
  return(d)
}
```

```{r}
d1 <- fna_1(paste0("./data/",files[1]))
Data2 <- d1

options(warn=2)
for (i in 2:length(files)) {
  di <- fna_2(paste0("./data/",files[i]))
  Data2 <- bind_rows(Data2,di)
  print(i)
}
```
```{r}
names(Data2)
myNames <- data.frame(names=names(Data2))
myNames %>% arrange(names)
```


The higher categories are:
Left column:
Total human exposures to Reason for exposure, give subcategory "Age:"
Unintentional to Intentional, give subcategory "RFE-Unintent:"
Intentional to Other, give subcategory "RFE-Intent:"
Other to end, give subcategory "RFE-Other:"

Right column: 
Management Site to Medical Outcome, give subcategory "MS"
Medical outcome to end, give subcategory "MO"


To achieve this, I will update the group function

```{r update gp}

gp <- function(p1Data){
  # group the data by column and height on the page
# keep the last entry of that column/height as the value
# assign the remaining entries for that column/height the name
  gd <- p1Data %>% 
    group_by(column,y) %>% 
    arrange(x, .by_group=TRUE) %>%
    mutate(type = ifelse(x==max(x) & x==min(x), "name", 
                         ifelse(x==max(x), "value", "name"))) 
  
    cd <- gd %>% 
      summarize(variable = paste(text[type=="name"], collapse=" "), 
                count=ifelse(is_empty(text[type=="value"])==FALSE, 
                             text[type=="value"],"0"), xmin=min(x))
      
 
y.age.min <- cd$y[cd$variable=="Total human exposures"]
y.age.max <- cd$y[cd$variable=="Reason for"]
y.un.min <- cd$y[cd$variable=="Unintentional"]
y.un.max <- cd$y[cd$variable=="Intentional"]
y.int.min <- cd$y[cd$variable=="Intentional"]

x.min <- min(cd$xmin[cd$column=="Left"])

y.int.max <- cd$y[cd$column=="Left" & cd$variable=="Other" & cd$xmin==x.min]
y.other.min <- cd$y[cd$column=="Left" & cd$variable=="Other" & cd$xmin==x.min]

y.ms.min <- cd$y[cd$variable=="Management"]
y.ms.max <- cd$y[cd$variable=="Medical"]
y.mo.min <- cd$y[cd$variable=="Medical"]

cd <- cd %>%
  mutate(variableSub=ifelse(column=="Left" & y > y.age.min & y < y.age.max, "Age:",
         ifelse(column=="Left" & y > y.un.min & y < y.un.max, "ReasonUn:",
         ifelse(column=="Left" & y > y.int.min & y < y.int.max, "ReasonInt:",
         ifelse(column=="Left" & y > y.other.min, "ReasonOther:",
         ifelse(column=="Right" & y > y.ms.min & y < y.ms.max, "MS:",
         ifelse(column=="Right" & y > y.mo.min, "MO:", "")))))))     
      
    
cd <- cd %>%
  filter(count != "Calls", count!="exposure", count!="Site",
           count!="Outcome",count!="Center",variable!="Maryland") %>%
  mutate(name=paste0(variableSub, variable))
  
  return(cd)
}


comb <- function(data, year, county){
  # create the data frame for this county/date
  myRow <- as.data.frame(t(as.numeric(gsub(",","",data$count))))
  names(myRow) <- data$name
  
  myRow$Year <- year
  myRow$County <- county
  return(myRow)
}
```

```{r}
fna_3 <- function(file){
  p1Data <- read(file)
  county <- ex(p1Data)[[1]]
  year <- ex(p1Data)[[2]]
  d <- mod(p1Data)
  d <- gp(d)
  d <- comb(d, year, county)
  return(d)
}
```

```{r}
d1 <- fna_3(paste0("./data/",files[1]))
Data3 <- d1

options(warn=2)
for (i in 2:length(files)) {
  di <- fna_3(paste0("./data/",files[i]))
  Data3 <- bind_rows(Data3,di)
}
```
```{r}
colnames(Data3)
```


1) WHY exist NA-sub-category? For "Allegany County Statistical Report 2016.pdf", the x.min of the left side doesn't equal to the x value for each subcatergory, so I update the x.min as x.min of "Total human exposures", since all subcategory are listed along the same x value. (change) 

```{r update gp}
gp <- function(p1Data){
  # group the data by column and height on the page
# keep the last entry of that column/height as the value
# assign the remaining entries for that column/height the name
  gd <- p1Data %>% 
    group_by(column,y) %>% 
    arrange(x, .by_group=TRUE) %>%
    mutate(type = ifelse(x==max(x) & x==min(x), "name", 
                         ifelse(x==max(x), "value", "name"))) 
  
    cd <- gd %>% 
      summarize(variable = paste(text[type=="name"], collapse=" "), 
                count=ifelse(is_empty(text[type=="value"])==FALSE, 
                             text[type=="value"],"0"), xmin=min(x))
      
 
y.age.min <- cd$y[cd$variable=="Total human exposures"]
y.age.max <- cd$y[cd$variable=="Reason for"]
y.un.min <- cd$y[cd$variable=="Unintentional"]
y.un.max <- cd$y[cd$variable=="Intentional"]
y.int.min <- cd$y[cd$variable=="Intentional"]

x.min <- cd$xmin[cd$variable=="Total human exposures"]

y.int.max <- cd$y[cd$column=="Left" & cd$variable=="Other" & cd$xmin==x.min]
y.other.min <- cd$y[cd$column=="Left" & cd$variable=="Other" & cd$xmin==x.min]

y.ms.min <- cd$y[cd$variable=="Management"]
y.ms.max <- cd$y[cd$variable=="Medical"]
y.mo.min <- cd$y[cd$variable=="Medical"]

cd <- cd %>%
  mutate(variableSub=ifelse(column=="Left" & y > y.age.min & y < y.age.max, "Age:",
         ifelse(column=="Left" & y > y.un.min & y < y.un.max, "ReasonUn:",
         ifelse(column=="Left" & y > y.int.min & y < y.int.max, "ReasonInt:",
         ifelse(column=="Left" & y > y.other.min, "ReasonOther:",
         ifelse(column=="Right" & y > y.ms.min & y < y.ms.max, "MS:",
         ifelse(column=="Right" & y > y.mo.min, "MO:", "")))))))     
      
    
cd <- cd %>%
  filter(count != "Calls", count!="exposure", count!="Site",
           count!="Outcome",count!="Center",variable!="Maryland") %>%
  mutate(name=paste0(variableSub, variable))
  
  return(cd)
}
```


```{r}
fna_4 <- function(file){
  p1Data <- read(file)
  county <- ex(p1Data)[[1]]
  year <- ex(p1Data)[[2]]
  d <- mod(p1Data)
  d <- gp(d)
  d <- comb(d, year, county)
  return(d)
}
```

```{r}
d1 <- fna_4(paste0("./data/",files[1]))
Data4 <- d1

options(warn=2)
for (i in 2:length(files)) {
  di <- fna_4(paste0("./data/",files[i]))
  Data4 <- bind_rows(Data4,di)
}
```
```{r}
colnames(Data4)
```

Now we can see there is no problem on subcategory!

Next, I will deal the whole dataframe. First define a function to find colnames with certain pattern

```{r}
#Data3$`Age:Adult`[is.na(Data3$`Age:<60 years`)]
str.dt <- function(pattern){
  temp <- colnames(Data4)[str_detect(colnames(Data4),pattern)]
  return(temp)
}
```

1) for subvar, I would like to sum all values in "Unknown" or "Other" as the new "unknown" col, and then delete orginal unknown value

```{r}
Data4$`Age:All.Unknown`= rowSums(Data4[str.dt("^Age:Unk")],na.rm = TRUE, dims=1)
Data4$`ReasonUn:All.Unknown`= rowSums(Data4[str.dt("ReasonUn:(Unk|Oth)")],na.rm = TRUE, dims=1)
Data4$`ReasonOther:All.Unknown`= rowSums(Data4[str.dt("ReasonOther:(Unk|Oth)")],na.rm = TRUE, dims=1)
Data4$`MS:All.Unknown`= rowSums(Data4[str.dt("MS:(Unk|Oth)")],na.rm = TRUE, dims=1)

```

2) for sub-Age, there are adult, >20-, >60, sum their value in to new 


```{r}
str.dt("Age:")
tempage <- Data4[c(str.dt("Age:"),"County")]
```

```{r}
Data4$`Age:>20 years`= rowSums(Data4[str.dt("Age:(Ad|20-|>60|<60)")],na.rm = TRUE, dims=1)
```

3)
```{r}
str.dt("ReasonUn:")
tempun <- Data4[c(str.dt("ReasonUn:"),"County","Year")]
```
```{r}
str.dt(".(Poi|poi)")
Data4$`ReasonUn:Botulism`= rowSums(Data4[str.dt(".(Poi|poi)")],na.rm = TRUE, dims=1)
str.dt(".(or Sti|or sti)")
Data4$`ReasonUn:Bite/Sting`= rowSums(Data4[str.dt(".(or Sti|or sti)")],na.rm = TRUE, dims=1)
```
Except these four variables, others have less NA, leave for later update

4) if we look at the sub-other, we can find there exist lots of NA's, how about sum them as a new column?

```{r}
tempoth <- Data4[c('Other',str.dt("ReasonOther:"),"County","Year")]
summary(tempoth)
```
```{r}
str.dt(".(With|Adv).")
Data4$`ReasonOther:Negative Reaction/Other`= rowSums(Data4[str.dt(".(With|Adv).")],na.rm = TRUE, dims=1)
```

put space after "/" in new variable as "/ " to avoid deleting it in later step

```{r}
str.dt(".(/Tamp|/tamp).")
Data4$`ReasonOther:Contamination/ Tampering`= rowSums(Data4[str.dt(".(/Tamp|/tamp).")],na.rm = TRUE, dims=1)
```

5) For sub-MS category:

only one NA for MS:Refused Referral, here I think we can replace that only NA as 0, leave this step to finish later, since we may need to repeat this step for other columns too.

```{r}
tempms <- Data4[c(str.dt("MS:"),"County","Year")]
summary(tempms)
```

6) No missing value for MO sub-category, good!

```{r}
tempmo <- Data4[c(str.dt("MO:"),"County","Year")]
summary(tempmo)
```

7) two wired vars 'Medical','ReasonUn:Outcome', which only has one value for "Calvert County, MD". This is caused by wired format of original pdf, just delete it

```{r}
summary(Data4[c('Medical','ReasonUn:Outcome')])
```

8) In all, we make change on 10 variables and now let's drop all old variables and save into a new dataframe
```{r}
drops <- c(str.dt("^Age:Unk"),str.dt("ReasonUn:Unk"),
           str.dt("ReasonOther:(Unk|Oth)"),str.dt("MS:(Unk|Oth)"),
           str.dt("Age:(Ad|20-|>60|<60)"),str.dt(".(Poi|poi)"),
           str.dt(".(or Sti|or sti)"), str.dt(".(With|Adv)."),
           str.dt(".(/Tamp|/tamp)."), "Medical",'ReasonUn:Outcome')

Data5 <- Data4[ , !(names(Data4) %in% drops)]
names(Data5)
```
9) change all NA as 0 or leave there? 

ReasonOther:Malicious has most NA:28

